
---
title: "Dairy Production"
format: 
  dashboard:
    nav-buttons: [github]
    github: https://github.com/WilliamBanda/Data-Dashboards-User-Perception
theme: [Cerulean, theme/custom.scss]
fig-width: 4
fig-asp: 0.9
params:
  month: 
  year: 
  threshold_diff: 0.01
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: load-packages
#| message: false
#| warning: false
#| include: false

options(repos = "https://cran.rstudio.com")

# Suppress messages and warnings globally
suppressPackageStartupMessages({
  install.packages("reshape2", quiet = TRUE)
  install.packages(c("writexl", "viridis", "shiny", "ggplot2"), quiet = TRUE)
  
  library(tidyverse)
  library(readr)
  library(dplyr)
  library(openxlsx)
  library(ggplot2)
  library(shiny)
  library(writexl)
  library(plotly)
  library(reshape2)
  library(readxl)
  library(scales)
  library(DT)
  library(gt)
  library(viridis)
  library(lubridate)
})

```

```{r}
#| label: load-data
#| message: false

Data <- read_excel("C:/Users/WilliamBanda1/Documents/MRP/tidy_data.xlsx")
Data <- Data %>%
  filter(format(as.Date(date), "%Y") %in% 2018:2023)
Data$year <- factor(year(Data$date))

```

```{r}
#| label: values
#| results: hide

# Calculate average milk produced from 2018-2023
avg_milk_produced <- Data %>%
  summarize(avg_milk_produced = mean(milk_produced, na.rm = TRUE)) %>%
  pull(avg_milk_produced) %>% round(digits = 1)

# Calculate average milk per cow from 2018-2023
avg_milk_per_cow <- Data %>%
  summarize(avg_milk_per_cow = mean(milk_per_cow, na.rm = TRUE)) %>%
  pull(avg_milk_per_cow) %>% round(digits = 1)

# Calculate average protein % from 2018-2023
avg_protein <- Data %>%
  summarize(avg_protein = mean(protein, na.rm = TRUE)) %>%
  pull(avg_protein) %>% round(digits = 1)

# Calculate average fat from 2018-2023 and round to one decimal point
avg_fat <- Data %>%
  summarize(avg_fat = round(mean(fat, na.rm = TRUE), 1)) %>%
  pull(avg_fat)

```
# About

## Row {.flow}
### column
::: card
::: card-header
About this Project and Data
:::
::: card-body
This project aims to address challenges of farm data management through organizing data and improving data awareness and accessibility. The aim is to improve accessibility, usability, compatibility, and convenience to end users or other stakeholders. 

The data used in this dashboard is from the Harper Adams Future Farm Dairy Unit. The data has dairy production and milk quality variables recorded on daily basis from 2018 to 2023.

[<i class="bi bi-github"></i> Source code](https://github.com/WilliamBanda/Dairy_Dashboard)

:::
:::

### column
::: card
::: card-header

Follow These Steps to Navigate this Dashboard 
:::
::: card-body
Step 1: Click on the Stats tab and take a glance of the types of graphs, colours and texts used.

Step 2: Try to interact with the graphs by hovering on top of one of the graphs with the cursor  and try to click on the extend button at the bottom right corner of the graph and hover through the graph.

Step 3: Click on the Data tab and take a glance of the data and the dictionary, try click the download data button or the download dictionary button.

Can you help answer the research questions?
:::
:::

# Stats
## Row {height=20%}

```{r}
#| content: valuebox
#| title: "Total Milk Produced (kg)"

list(
  color = "purple",
  value = avg_milk_produced
)
```

```{r}
#| content: valuebox
#| title: "Average Milk per Day (kg)"

list(
  color = "pink",
  value = avg_milk_per_cow
)
```

```{r}
#| content: valuebox
#| title: "Milk Quality (Protein %)"

list(
  color = "cyan",
  value = avg_protein
)
```

```{r}
#| component: valuebox
#| title: "Fat % Average"

list(
  color = "red",
  value = avg_fat
)
```

## Row {height=50%}

```{r}
#| title: "Daily Average Milk per Cow per Year"

# Extract year from the date column
Data$year <- lubridate::year(Data$date)

# Define a color palette
palette <- c("blue", "red", "green", "purple", "orange", "brown")

# Determine the maximum and minimum values for the y-axis
max_milk_per_cow <- max(Data$milk_per_cow, na.rm = TRUE)
min_milk_per_cow <- min(Data$milk_per_cow, na.rm = TRUE)

# Create the boxplot
p <- Data %>%
  filter(milk_per_cow < 299) %>%  # Adjust the filter condition as needed
  ggplot(aes(x = factor(year), y = milk_per_cow, fill = factor(year))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 1, lwd = 0.1) +  # Add outliers and set line width
  geom_jitter(width = 0.2, alpha = 0.5) +  # Add data points with jitter
  geom_hline(yintercept = max_milk_per_cow, linetype = "dashed", color = "red") +  # Line at max value
  geom_hline(yintercept = min_milk_per_cow, linetype = "dashed", color = "blue") +  # Line at min value
  labs(x = "Year", y = "Milk per Cow") +  # Update the y-axis label
  scale_fill_manual(values = palette) +  # Apply the color palette
  theme_bw() +
  theme(legend.position = "none") +  # Remove the legend
  scale_y_continuous(limits = c(min_milk_per_cow - 0.1 * (max_milk_per_cow - min_milk_per_cow),
                                max_milk_per_cow + 0.1 * (max_milk_per_cow - min_milk_per_cow)),
                     expand = c(0, 0))  # Adjust y-axis limits to increase box size

print(p)

```

### Column

```{r}
#| title: "Yearly Average Protein Percentage"

# Extract year from the date column
Data$year <- lubridate::year(Data$date)

# Define a color palette
palette <- c("blue", "red", "green", "purple", "orange", "brown")

# Determine the maximum and minimum values for the y-axis
max_protein <- max(Data$protein, na.rm = TRUE)
min_protein <- min(Data$protein, na.rm = TRUE)

# Set limits for y-axis
lower_limit <- 2.5
upper_limit <- 5

# Create the boxplot
p <- Data %>%
  filter(protein <= 4) %>%  # Filter data points up to 4
  ggplot(aes(x = factor(year), y = protein, fill = factor(year))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 1, lwd = 0.1) +  # Add outliers and set line width
  geom_jitter(width = 0.2, alpha = 0.5) +  # Add data points with jitter
  labs(x = "Year", y = "Protein %") +
  scale_fill_manual(values = palette) +  # Apply the color palette
  theme_bw() +
  theme(legend.position = "none") +  # Remove the legend
  scale_y_continuous(limits = c(lower_limit, upper_limit),
                     expand = c(0, 0)) +  # Set y-axis limits
  geom_hline(yintercept = max_protein, linetype = "dashed", color = "red") +  # Add maximum value line
  geom_hline(yintercept = min_protein, linetype = "dashed", color = "blue")   # Add minimum value line

print(p)


```
### Column
```{r}
#| title: Yearly Average Fat Percentage

# Extract year from the date column
Data$year <- lubridate::year(Data$date)

# Define a color palette
palette <- c("blue", "red", "green", "purple", "orange", "brown")

# Determine the maximum and minimum values for the y-axis
max_fat <- max(Data$fat, na.rm = TRUE)
min_fat <- min(Data$fat, na.rm = TRUE)

# Create the boxplot
p <- Data %>%
  filter(fat < 299) %>%
  ggplot(aes(x = factor(year), y = fat, fill = factor(year))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 1, lwd = 0.1) +  # Add outliers and set line width
  geom_jitter(width = 0.2, alpha = 0.5) +  # Add data points with jitter
  labs(x = "Year", y = "Fat %") +
  scale_fill_manual(values = palette) +  # Apply the color palette
  theme_bw() +
  theme(legend.position = "none") +  # Remove the legend
  scale_y_continuous(limits = c(min_fat - 0.1 * (max_fat - min_fat), max_fat + 0.1 * (max_fat - min_fat)),
                     expand = c(0, 0)) +  # Adjust y-axis limits to increase box size
  geom_hline(yintercept = max_fat, linetype = "dashed", color = "red") +  # Add maximum value line
  geom_hline(yintercept = min_fat, linetype = "dashed", color = "blue")   # Add minimum value line

print(p)

```
# Data

```{r}

# Load required libraries
# Sample data
Data <- data.frame(
  Date = Sys.Date() - 1:100,
  CowsInMilk = sample(100:200, 100, replace = TRUE),
  CowsInTank = sample(100:200, 100, replace = TRUE),
  MilkProduced = sample(500:1000, 100, replace = TRUE),
  MilkPerCow = runif(100, 20, 30),
  StrawYard = sample(50:100, 100, replace = TRUE),
  Highs = sample(10:20, 100, replace = TRUE),
  Trial1 = sample(5:15, 100, replace = TRUE),
  Trial2 = sample(5:15, 100, replace = TRUE),
  Low = sample(1:10, 100, replace = TRUE),
  Hscc = sample(150:200, 100, replace = TRUE),
  Scc = sample(200:300, 100, replace = TRUE),
  Bs = runif(100, 3, 5),
  Fat = runif(100, 3, 5),
  Protein = runif(100, 3, 5)
)

# Create the interactive table
table <- datatable(
  Data,
  colnames = c("ID", "Date", "Cows in Milk", "Cows in Tank", "Milk Produced", "Milk per Cow", "Straw Yard", "Highs", "Trial 1", "Trial 2", "Low", "Hscc", "Scc", "Bs", "Fat", "Protein"),
  options = list(
    dom = "Bfrtip",
    buttons = list(
      list(
        extend = 'excel',
        text = 'Download Data',
        title = 'Dairy Production Data'
      )
    ),
    paging = TRUE,
    lengthMenu = c(10, 50, 100, 150, 200, 500, 1000, 1500),
    pageLength = 10,
    columnDefs = list(list(className = 'dt-center', targets = "_all")),
    scrollY = FALSE,
    initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().container()).find('th, td').css({'color': 'black'});",  # Set text color to black
      "$('.dt-button').css({'background-color': '#007a78', 'color': 'white'});",  # Style buttons
      "}"
    )
  ),
  extensions = "Buttons"
) %>%
  formatRound(columns = c("Fat", "Bs", "Protein", "MilkPerCow"), digits = 2)

# Format the date column
table <- formatDate(
  table,
  columns = 1,
  method = "toDateString"
)

# Display the interactive table
table


```

```{r}
#| title: Dictionary

# Read the Excel file and load the data from the second sheet
Dictionary <- read_excel("C:/Users/WilliamBanda1/Documents/MRP/tidy_data.xlsx", sheet = 2)

# Create the DataTable with the specified options
datatable(
  Dictionary,
  options = list(
    dom = "Bfrtip",
    buttons = list(
      list(
        extend = 'excel',
        text = 'Download Dictionary',
        title = 'Dairy Data Dictionary'
      )
    ),
    paging = TRUE,
    pageLength = 10,
    columnDefs = list(list(className = 'dt-center', targets = "_all")),
    scrollY = FALSE,
    initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().container()).find('th, td').css({'color': 'black'});",  # Set text color to black
      "$('.dt-button').css({'background-color': '#007a78', 'color': 'white'});",  # Style buttons
      "}"
    )
  )
)


```



